Gammapy: A Python package for gamma-ray astronomy

### 2023-01-27 src/ms-review.tex
Smoothen the introduction a bit

% \documentclass[]{aa}
\documentclass[traditabstract, longauth]{aa}

@@

% Note: not sure if we want to use that ... doesn't look too pretty
\newcommand{\astropy}{Astropy\xspace}
\newcommand{\gammapy}{Gammapy\xspace}
\newcommand{\scipy}{Scipy\xspace}
\newcommand{\numpy}{Numpy\xspace}
\newcommand{\iminuit}{IMinuit\xspace}
\newcommand{\sherpa}{Sherpa\xspace}
\newcommand{\agnpy}{Agnpy\xspace}

@@

%% \gammaray astronomy
%% CN: I think we should use the expression Water Cherenkov Detectors and use the
%% abbreviation (WCD) all along the text, later we use "particle samplers"
\gammaray astronomy is a rather young field of research. The \gammaray
range of the electromagnetic spectrum provides us insights into the
most energetic processes in the universe such as those accelerating particles in the surroundings of
black holes, and remnants of supernova explosions. As in other
branches of astronomy, \gammarays can be observed by both
satellite as well as ground based instruments.
Ground-based instruments use the Earth's atmosphere as a particle detector.
Very-high-energy (VHE) cosmic \gammarays interact in the atmosphere and
create a large shower of secondary particles that can be observed from the ground.
Ground-based \gammaray astronomy relies on these extensive air showers to detect the
primary \gammaray photons and infer their incident direction and energy.
VHE \gammaray astronomy covers the energy range from fews tens of ${\rm GeV}$ up to the ${\rm PeV}$.
There are two main categories of ground-based instruments: 
Imaging Atmospheric Cherenkov Telescopes (IACT) and 
Water Cherenkov Detectors (WCD). IACTs obtain images of the atmospheric showers
by detecting the Cherenkov radiation emitted by the cascading charged particles and
use these images to reconstruct the properties of the incident particle.
Those instruments have a limited field of view (FoV) and duty cycle, but
good energy and angular resolution.

@@

Water Cherenkov observatories detect particles directly from the tail of the
shower when it reaches the ground. These instruments have a very
large FoV large duty-cycle but higher energy threshold and
usually have lower signal to noise ratios compared to IACTs \citep{2015CRPhy..16..610D}.

@@

%% Context: develoment of open software
In the past decade Python has become extremely popular as a scientific programming language,
in particular, in the field of data sciences. This success is
mostly attributed to the simple and easy to learn syntax, the ability to act as
a "glue" language between different programming languages and last but not least
the rich ecosystem of packages and its open and supportive community, from
core numerical analysis packages such as \numpy~\citep{numpy} and
\scipy~\citep{2020SciPy-NMeth}, visualization libraries, e.g. Matplotlib \citep{matplotlib},
to optimization packages such as \iminuit \citep{iminuit}.

The \astropy~\citep{astropy} project was created in 2012 to build a community-developed
Python package for astronomical research. Among other things, it offers functionalities for
transforming and representing astronomical coordinates, manipulating physical quantities
and their units as well as reading and writing FITS files.

%% \gammapy: concept and goals
%%% Question: starting date? Is is relevant, what is the real start?
The \gammapy project started based on the same idea as \astropy with the objective of building a flexible and
efficient software library for very high-energy \gammaray data analysis \citep{gammapy_2015}.
This is possible thanks to the definition of a standardised common data format such as the \gadf, and thanks to
an open development with a broad community of contributors. The relation of \gammapy to the
different instruments and the \gadf data format is illustrated in Figure~\ref{fig:big_picture}.

%% CN: what about mentioning the joint Crab? or is it mentioned later?
The \hess collaboration released a limited test dataset (about 50 hours of
observations taken between 2004 and 2008) based  on the \gadf DL3 format \citep{HESS_DR1}.
This data release served as a basis for validation of open analysis tools,
including \gammapy \cite[see e.g.][]{Mohrmann2019}.

>>>>

% \documentclass[]{aa}
\documentclass[longauth]{aa}

@@

% Note: not sure if we want to use that ... doesn't look too pretty
\newcommand{\astropy}{Astropy\xspace}
\newcommand{\gammapy}{Gammapy\xspace}
\newcommand{\scipy}{Scipy\xspace}
\newcommand{\numpy}{Numpy\xspace}
\newcommand{\iminuit}{IMinuit\xspace}
\newcommand{\sherpa}{Sherpa\xspace}
\newcommand{\agnpy}{Agnpy\xspace}
\newcommand{\matplotlib}{Matplotlib\xspace}

@@

%% \gammaray astronomy
%% CN: I think we should use the expression Water Cherenkov Detectors and use the
%% abbreviation (WCD) all along the text, later we use "particle samplers"
\gammaray astronomy is a rather young field of research. The \gammaray
range of the electromagnetic spectrum provides us insights into the
most energetic processes in the universe such as those accelerating particles in the surroundings of
black holes, and remnants of supernova explosions. As in other
branches of astronomy, \gammarays can be observed by both
satellite as well as ground based instruments.
Ground-based instruments use the Earth's atmosphere as a particle detector.
Very-high-energy (VHE) cosmic \gammarays interact in the atmosphere and
create a large shower of secondary particles that can be observed from the ground.
Ground-based \gammaray astronomy relies on these extensive air showers to detect the
primary \gammaray photons and infer their incident direction and energy.
VHE \gammaray astronomy covers the energy range from fews tens of ${\rm GeV}$ up to the ${\rm PeV}$.
There are two main categories of ground-based instruments:

\textit{Imaging Atmospheric Cherenkov Telescopes (IACT)} obtain images of the atmospheric showers
by detecting the Cherenkov radiation emitted by the cascading charged particles and
use these images to reconstruct the properties of the incident particle.
Those instruments have a limited field of view (FoV) and duty cycle, but
good energy and angular resolution.

\textit{Water Cherenkov Detectors (WCD)} detect particles directly from the tail of the
shower when it reaches the ground. These instruments have a very
large FoV large duty-cycle but higher energy threshold and
usually have lower signal to noise ratios compared to IACTs \citep{2015CRPhy..16..610D}.

@@

@@

%% Context: development of open software
In the past decade observing the \gammaray sky has transitioned from a niche in the field of
particle physics to an established branch of astronomy, completing the view of the
sky in high energies. At the same time \textit{Python}
has become extremely popular as a scientific programming language,
in particular, in the field of data sciences. This success is
mostly attributed to the simple and easy to learn syntax, the ability to act as
a "glue" language between different programming languages and last but not least
the rich ecosystem of packages and its open and supportive community \citep{Momcheva2015}.

In the sub-field of astronomy, it was the \astropy project \citep{astropy} that 
was created in 2012 to build a community-developed core Python package for astronomy.
It offers basic functionalities that astronomers of many fields needs, such as representing
and transforming astronomical coordinates, manipulating physical quantities includinf units
as well as reading and writing FITS files.

%% \gammapy: concept and goals
%%% Question: starting date? Is is relevant, what is the real start?
The \gammapy project was started following the idea of \astropy: the objective of building a common
software library for very high-energy \gammaray data analysis \citep{gammapy_2015}. 
The core of the idea is illustrated in Figure~\ref{fig:big_picture}. Various \gammaray instruments
export their data to a standardised common data format the \gadf. This data can then be combined and
analysed using a single common software library. This means that the \gammapy package is not a part of
any instrument, but an independent community developed software project. The \gammapy package is
is built on the scientific Python ecosystem: it uses \numpy~\citep{numpy} for n-dimensional data
structures, \scipy~\citep{2020SciPy-NMeth} for numerical algorithms, \astropy~\citep{astropy} for
astronomy-specific functionality, and \matplotlib~\citep{matplotlib} for visualization.

%% CN: what about mentioning the joint Crab? or is it mentioned later?
With the public availability of the \gadf format sepcifications and the 
\gammapy package, some experiments started to make limited subsets of
their \gammaray data publicly available for testing and validating
\gammapy. For example, the \hess collaboration released a limited test
dataset (about 50 hours of observations taken between 2004 and 2008)
based  n the \gadf DL3 format \citep{HESS_DR1}. This data release served
as a basis for validation of open analysis tools, including \gammapy 
\cite[see e.g.][]{Mohrmann2019}. The \hawc collaboration also released
a limited test dataset of the Crab Nebula, which was used to validate
the \gammapy package in \cite{Olivera2022}.
